Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-21 code-tests.tex
Added gravity and self-similar tests

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Point source gravity test}
\label{sec.test.gravitypointsource}
\red{(Greg)}
This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Self-Similar infall test}
\label{sec.tests.infall}
\red{(Greg)}
Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
solution, and tests gravity + hydro.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Point source gravity test}
\label{sec.test.gravitypointsource}

This test is a simple test of the ability of the Poisson solver to correctly reproduce the acceleration around a point source without any dynamics.  We place a single particle in the center of the unit domain with root grid dimensions $32^3$ and add a static AMR grid with refinement factor two that covers the central $1/8^3$ of the domain (i.e. so the refined grid is only $8^3$).  We then place 5000 particles throughout the domain distributed uniformly in angle and with the logarithm of the radius (from the central point).  The measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As is typical for Particle-Mesh based methods, the errors decrease at large distances and peak on roughly the grid scale.  The accelerations are relatively smooth across the top grid/fine grid boundary with the most noticeable impact being a drop in the amplitude of tangential errors on the fine grid.  The force law rolls over at about 1.6 fine cell widths (equivalent to 0.8 root grid widths in this figure), consistent with the CIC-deposition and interpolation used.

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/GravityTest.eps}
\caption{The radial and tangential accelerations measured around a point source placed in a static AMR grid.  The solid line shows the analytic radial profile.}
\label{fig.gravitytest}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Self-Similar infall test}
\label{sec.tests.infall}

The next test problem is based on the self-similar solutions to the cosmological spherical infall problem found by \citet{Bertschinger1985}.  It features a small density perturbation in a homogeneous $\Omega = 1$ universe and is a strong test of the cosmological evolution, gravity, and hydrodynamic portions of the code.  It is close analogue to cosmological cluster formation and has an analytic solution.

For initial conditions, we adopt a $32^3$ top grid with a single, initial sub grid (covering $1/8^3$ of the domain) with a refinement factor of 2.  An overdensity $\delta = 40$ is placed in a single cell near the center of the domain.  We begin at $z=199$ and evolve to $z=0$, which is a sufficient time for the evolution to largely forget its initial conditions and approach the self-similar result.  An over density refinement criteria ($\delta_{\rm crit} = 1.1$ on the top grid) is used to add additional grids, going up to 5 additional levels beyond the root grid.  We use the PPM solver without radiative cooling; only baryons are used for this problem.  An ideal gas law with $\gamma = 5/3$ is adopted.

In Figure~\ref{fig.sphericalinfall}, we show the results, scaled according to the dimensionless variables as defined in equations (2.9) and (3.2) of \citet{Bertschinger1985}.  This demonstrates that we can quickly and easily obtain a good solution with only a fairly modest initial grid.  The shock is sharply resolved and the asymptotic profiles as small $\lambda$ are recovered.  At very small values of $\lambda$, the initial conditions have not been fully forgotten and the self-similar solution is not recovered -- this can be seen most clearly in the dimensionless mass; however, this is simply because of the limited amount of time for which we evolved the solution.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/SphericalInfall.eps}
\caption{The results of the self-similar spherical infall test.  The panels show, as a function of the dimensionless radius $\lambda$, the dimensionless density $D$ (top left), velocity $V$ (top right), pressure $P$ (bottom left), and enclose mass $M$ (bottom right).  In each case we show azimuthally averaged profiles from the simulation as circles and the analytic solution for for $\gamma = 5/3$ as solid lines.}
\label{fig.sphericalinfall}
\end{center}
\end{figure}

### 2013-05-21 numerical-starformation.tex
Added gravity and self-similar tests

Due to the computationally unfeasible number of stars in a galaxy
(10^11) and the lack of detailed understanding of star formation, a
number of phenomenological star formation models are included in
\enzo.
The \enzo\ code includes approximately ten models for star formation
and feedback.  Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but may include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell, with the same
momentum vector as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.

>>>>

Due to the computationally unfeasible number of stars in a galaxy
($10^{11}$) and the lack of detailed understanding of star formation, a
number of phenomenological star formation models are included in
\enzo.
The \enzo\ code includes approximately ten models for star formation
and feedback.  Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but may include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell, with the same
momentum vector as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.
